Interpreting Random Forest Predictions for Firearm Identification Using LIME

### 2018-11-01 paper/paper.Rmd
moved all data steps to paper from app and cleaned up the code - also added mark to app

  # Clean the Hamby 224 set 1 data
  hamby224_set1_cleaned <- hamby224_set1 %>%
    select(-bullet_score, -land1, -land2, -aligned, -striae, -features) %>%
    rename(bullet1 = bulletA,
           bullet2 = bulletB, 
           land1 = landA,
           land2 = landB) %>%
    mutate(study = factor("Hamby_224"), 
           set = factor("1"),
           bullet1 = factor(bullet1),
           bullet2 = factor(bullet2),
           land1 = factor(land1),
           land2 = factor(land2)) %>%
    select(study, set, bullet1:land2, rf_features, rfscore, samesource)

  # Clean the Hamby 224 set 11 data
  hamby224_set11_cleaned <- hamby224_set11 %>%
    select(-bullet_score, -land1, -land2, -aligned, -striae, -features) %>%
    rename(bullet1 = bulletA,
           bullet2 = bulletB, 
           land1 = landA,
           land2 = landB) %>%
    mutate(study = factor("Hamby_224"), 
           set = factor("11"),
           bullet1 = recode(factor(bullet1), 
                            "Bullet 1" = "1", 
                            "Bullet 2" = "2", 
                            "Bullet I" = "I"),
           bullet2 = recode(factor(bullet2), 
                            "Bullet 1" = "1", 
                            "Bullet 2" = "2", 
                            "Bullet I" = "I"),
           land1 = recode(factor(land1), 
                          "Land 1" = "1", "Land 2" = "2", "Land 3" = "3", 
                          "Land 4" = "4", "Land 5" = "5", "Land 6" = "6"),
           land2 = recode(factor(land2), 
                          "Land 1" = "1", "Land 2" = "2", "Land 3" = "3", 
                          "Land 4" = "4", "Land 5" = "5", "Land 6" = "6")) %>%
    select(study, set, bullet1:land2, rf_features, rfscore, samesource)

  # Join the two cleaned Hamby 224 sets into one testing set
  hamby224_test <- suppressWarnings(bind_rows(hamby224_set1_cleaned,
                                              hamby224_set11_cleaned)) %>%
    mutate(case = as.character(1:length(study))) %>%
    select(case, study:samesource)

@@

  # Apply the explain function from the lime package (with a seed set)
  set.seed(84902)
  hamby224_explain <- lime::explain(hamby224_test %>% select(rf_features), 
                                    hamby224_lime,
                                    labels = "FALSE",
                                    n_features = 3)

@@

```{r echo = FALSE}
# Function to use for creating bin labels in the test_explain dataset
bin_labeller <- function(bin_cuts, feature, feature_value){

@@

  # Determine which bin the case falls in
  if(feature_value <= feature_bin_cuts$Q25){
    feature_bin <- paste("lower", feature, "bin")
  } else if (feature_bin_cuts$Q25 < feature_value & feature_value <= feature_bin_cuts$Q50){
    feature_bin <- paste("lower middle", feature, "bin")
  } else if (feature_bin_cuts$Q50 < feature_value & feature_value <= feature_bin_cuts$Q75){
    feature_bin <- paste("upper middle", feature, "bin")
  } else if (feature_bin_cuts$Q75 < feature_value){
    feature_bin <- paste("upper", feature, "bin")
  }

@@

  # Join the data and the explanations and edit and add additional variables
  hamby224_test_explain <- hamby224_test %>%
    mutate(case = as.character(case)) %>%
    full_join(hamby224_explain, by = "case") %>%
    mutate(case = factor(case),
           set = factor(set),
           bullet1 = factor(bullet1),
           bullet2 = factor(bullet2),
           land1 = factor(land1),
           land2 = factor(land2)) %>%
    mutate(mypred = rep(predict(rtrees, hamby224_test %>% select(rf_features), 
                                type = "prob")[,2], each = 3)) %>%
    select(case:rfscore, mypred, samesource:prediction) %>%
    mutate(feature_desc = factor(feature_desc),
           feature = factor(feature)) %>%
    mutate(feature_number = readr::parse_number(feature_desc),
           strictly_less = FALSE) %>%
    mutate(feature_bin = mapply(bin_labeller, feature = feature, feature_value = feature_value))

@@

```{r}
hamby224_test_explain %>%
  filter(set == "1") %>%
  select(case, bullet1:land2, rfscore) %>%
  distinct() %>%
  ggplot(aes(x = land1, y = land2)) +
  geom_tile(aes(fill = rfscore)) +
  facet_grid(bullet1 ~ bullet2) +
  theme_bw() +
  scale_fill_gradient2(low = "grey", high = "orange", midpoint = 0.5)
```

>>>>

  # Clean the Hamby 224 set 1 data
  hamby224_set1_cleaned <- hamby224_set1 %>%
    select(-bullet_score, -land1, -land2, -aligned, -striae, -features) %>%
    rename(bullet1 = bulletA,
           bullet2 = bulletB, 
           land1 = landA,
           land2 = landB) %>%
    mutate(study = factor("Hamby 224"), 
           set = factor("Set 1"),
           bullet1 = recode(factor(bullet1), "1" = "Known 1", "2" = "Known 2", "Q" = "Questioned"),
           bullet2 = recode(factor(bullet2), "1" = "Known 1", "2" = "Known 2", "Q" = "Questioned"),
           land1 = recode(factor(land1), "1" = "Land 1", "2" = "Land 2", "3" = "Land 3", 
                          "4" = "Land 4", "5" = "Land 5", "6" = "Land 6"),
           land2 = recode(factor(land2), "1" = "Land 1", "2" = "Land 2", "3" = "Land 3", 
                          "4" = "Land 4", "5" = "Land 5", "6" = "Land 6")) %>%
    select(study, set, bullet1:land2, rf_features, rfscore, samesource)

  # Clean the Hamby 224 set 11 data
  hamby224_set11_cleaned <- hamby224_set11 %>%
    select(-bullet_score, -land1, -land2, -aligned, -striae, -features) %>%
    rename(bullet1 = bulletA,
           bullet2 = bulletB, 
           land1 = landA,
           land2 = landB) %>%
    mutate(study = factor("Hamby 224"), 
           set = factor("Set 11"),
           bullet1 = recode(factor(bullet1), "Bullet 1" = "Known 1", "Bullet 2" = "Known 2", 
                            "Bullet I" = "Questioned"),
           bullet2 = recode(factor(bullet2), "Bullet 1" = "Known 1", "Bullet 2" = "Known 2", 
                            "Bullet I" = "Questioned")) %>%
    select(study, set, bullet1:land2, rf_features, rfscore, samesource)

  # Create a dataset with all combinations of lands and bullets comparisons for each set
  combinations <- data.frame(set = factor(rep(c("Set 1", "Set 11"), each = 324)),
                      expand.grid(land1 = factor(c("Land 1", "Land 2", "Land 3", 
                                                   "Land 4", "Land 5", "Land 6")),
                                  land2 = factor(c("Land 1", "Land 2", "Land 3", 
                                                   "Land 4", "Land 5", "Land 6")),
                                  bullet1 = factor(c("Known 1", "Known 2", "Questioned")),
                                  bullet2 = factor(c("Known 1", "Known 2", "Questioned"))))

  # Join the two cleaned Hamby 224 sets into one testing set
  hamby224_test <- suppressWarnings(bind_rows(hamby224_set1_cleaned,
                                              hamby224_set11_cleaned)) %>%
    mutate(set = factor(set),
           bullet1 = factor(bullet1),
           bullet2 = factor(bullet2),
           land1 = factor(land1),
           land2 = factor(land2)) %>%
    right_join(combinations, by = c("set", "land1", "land2", "bullet1", "bullet2")) %>%
    filter(!(bullet1 == "Questioned" & bullet2 == "Known 1"),
           !(bullet1 == "Questioned" & bullet2 == "Known 2"),
           !(bullet1 == "Known 2" & bullet2 == "Known 1")) %>%
    arrange(rfscore) %>%
    mutate(case = factor(1:length(study))) %>%
    select(case, study:samesource)

@@

  # Apply the explain function from the lime package (with a seed set)
  set.seed(84902)
  hamby224_explain <- lime::explain(hamby224_test %>% 
                                      arrange(case) %>% 
                                      select(rf_features) %>% 
                                      na.omit(), 
                                    hamby224_lime,
                                    labels = "FALSE",
                                    n_features = 3)

@@

```{r echo = FALSE}
# Function to use for creating bin labels in the test_explain dataset
bin_labeller <- function(feature, feature_value){

  if (is.na(feature)) {

    # Set feature_bin to NA if feature is NA
    feature_bin <- NA

@@

    # Determine which bin the case falls in
    if(feature_value <= feature_bin_cuts$Q25){
      feature_bin <- paste(feature, "(lower bin)")
    } else if (feature_bin_cuts$Q25 < feature_value & feature_value <= feature_bin_cuts$Q50){
      feature_bin <- paste(feature, "(lower middle bin)")
    } else if (feature_bin_cuts$Q50 < feature_value & feature_value <= feature_bin_cuts$Q75){
      feature_bin <- paste(feature, "(upper middle bin)")
    } else if (feature_bin_cuts$Q75 < feature_value){
      feature_bin <- paste(feature, "(upper bin)")
    }

  }

@@

  # Join the data and the explanations and edit and add additional variables
  hamby224_test_explain <- hamby224_test %>%
    mutate(case = as.character(case)) %>%
    full_join(hamby224_explain, by = "case") %>%
    mutate(case = factor(case),
           feature_desc = factor(feature_desc),
           feature = factor(feature),
           feature_bin = mapply(bin_labeller, feature = feature, feature_value = feature_value)) %>%
    mutate(feature_number = readr::parse_number(feature_desc),
           strictly_less = FALSE)

@@

```{r}
hamby224_test_explain %>%
      filter(set == "Set 1") %>%
      mutate(rfscore = round(rfscore, 3)) %>%
      select(case, bullet1, bullet2, land1, land2, rfscore) %>%
      distinct() %>%
      ggplot(aes(x = land1, y = land2, label = bullet1, label2 = bullet2,
                 text = paste('Bullets Compared: ', bullet1, "-", land1, 
                              "vs", bullet2, "-", land2,
                              '\nRandom Forest Score: ', 
                              ifelse(is.na(rfscore), "Missing due to tank rash", rfscore)))) +
      geom_tile(aes(fill = rfscore)) +
      facet_grid(bullet2 ~ bullet1, scales = "free") +
      theme_minimal() +
      scale_fill_gradient2(low = "darkgrey", high = "darkorange", midpoint = 0.5) +
      theme(axis.text.x = element_text(angle = 45, hjust = 1),
            strip.text.x = element_blank(),
            strip.text.y = element_blank()) +
      labs(x = "Known 1                            Known 2                            Questioned", 
           y = "Known 1                            Known 2                            Questioned", 
           fill = "RF Score")
```
